---
title: "Gov 50 Final Project"
author: "Ananya Choudhary"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Introduction of the project

Is female representation in the parliament of a country a direct result of their access to education?
The direct correlation between women's representation in parliamentary bodies and their access to education is a well-documented and essential aspect of gender equality. Access to education equips women with the knowledge, skills, and confidence necessary to participate actively in the political sphere. Education can empower women to understand their rights, engage in public discourse, and navigate complex policy issues effectively. As women gain access to quality education, they are more likely to break through traditional gender roles and stereotypes, ultimately leading to increased female participation in political leadership positions, which is vital for a more equitable and representative democracy. My hypothesis is that as the access to education for women increases, their political representation in the parliament also increases. My sample consists of two datasets from the United Nations Organisation about all the member countries' seats held by women in the parliament and the gender ratio in access to education. For this project I have focused on South Asian countries namely India, Pakistan, Nepal, Bangladesh and Sri Lanka only. The goal is to determine whether there is a connection between access to education and political representation in the parliament in the context of women.  

A reason why this project focuses on South Asian countries is because these are the set of countries with varying levels of all the factors that could potentially affect the connection that has been tried to achieve here. GDP of the country, equitable and easy access to the internet, increasing or decreasing crime rates, reservation quota for women in parliament or even schools, government expenditure on schools and education are all examples of factors that impact the relationship between access to education and political representation in the parliament. This is problem of causality as all these varying levels serve as confounders to the research question. Therefore, to combat this issues and also show the impact such factors have on this relationship, this project draws a comparative analysis between the relationship of the two primary factors once without the confounders and then with the confounders with the help of the concept of regression. 



##Data Section

#The first dataset on seats held by women in parliament is loaded below. It has further been filtered to only South Asian countries.

Description of variables is as follows:

region_country_area = country code as per the UN
country = name of the country
year = year when the data was recorded
series = statement explaining the "year" variable
last_election_date = date when the last election was held
value = percentage value of the seats held by women in parliament

This project only focuses on the usage of the country, year, and value variable. 

```{r}
library(tidyverse)
library(janitor)
df <- read_csv("SYB65_317_202209_Seats held by women in Parliament.csv", skip = 1) |>
  janitor::clean_names() |> 
  rename(country = x2) 

southasian_countries <- df |>
  filter(country == "India"| country =="Pakistan"|country =="Sri Lanka"|country =="Bangladesh"|country=="Nepal")

ggplot(data = southasian_countries, mapping=aes(x=year, y=value, color = country, fill = country)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + facet_wrap(~country) + labs(x = "year", y = "percentage of women in parliament", title = "Relationship between representation of women in parliament with increasing years") 

```
##Discussion of the plot

This plot displays the relationship between the representation of women in parliament in percentage and time in years in 5 different South Asian countries. Through this visualization it is evident that percentage representation of women in parliament of Bangladesh, India and Nepal has increased as years have passed. However, Pakistan and Sri Lanka do not show the same relationship. It is striking to note that while Sri Lanka has shown the slightest increase in women representation in the parliament, Pakistan has shown a decrease in the same over the years. Also, it is imperative to note that aside from India, that has shown a consistent growth in the representation of women in parliament over the years, Bangladesh and Nepal show a rather clustered scatterplot with the increment in years.



#The second dataset on ratio between girls and boys in primary, secondary and upper secondary is loaded below. It has further been filtered to only South Asian countries.

Description of variables is as follows:

region_country_area = country code as per the UN
country = name of the country
year = year when the data was recorded
series = breakdown of levels of education into primary, secondary and upper secondary
value = value of the ratio between girls to boys

```{r}
daf <- read_csv("SYB65_319_202209_Ratio of girls to boys in education.csv", skip = 1) |>
 janitor::clean_names() |> 
  rename(country = x2) 

southasian_countries_edu <- daf |>
  filter(country == "India"| country =="Pakistan"|country =="Sri Lanka"|country =="Bangladesh"|country=="Nepal")

ggplot(data = southasian_countries_edu, mapping=aes(x=year, y=value, color = series, fill = series)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~country) + theme_minimal() + theme(legend.position="bottom") +  labs(x = "year", y = "value of ratio between girls and boys", title = "Relationship between ratio of girls and boys in different levels of education with increasing years") 
```
##Discussion of this plot

This plot displays the relationship between the ratio of girls and boys in different levels of education and time in years in 5 different South Asian countries. Through this visualization it is evident that ratio of girls and boys in different levels of education in India, Pakistan and Nepal has increased as years have passed. However, Pakistan and Sri Lanka do not show the same relationship. It is striking to note that Sri Lanka has shown the slightest increase in ratio in primary education but a sharp decrease in secondary and upper secondary levels of education. Bangladesh observes a similar trend but in reverse. There is only a slight increase in ratio of girls and boys in primary level of education but a decent increase in secondary and upper secondary levels of education. 


##loading confounder datasets

```{r}
GDP <- read_csv("GDP per capita.csv", skip = 1) |>
 janitor::clean_names() |> 
  rename(country = x2) |>
   filter(country == "India"| country =="Pakistan"|country =="Sri Lanka"|country =="Bangladesh"|country=="Nepal",
          series == "GDP per capita (US dollars)") |>
  rename(gdp_per_capita = "value") |>
  select(country, gdp_per_capita, year )
GDP

internet_usage <- read.csv("Internet usage.csv", skip = 1)|>
   filter(X == "India"| X =="Pakistan"|X =="Sri Lanka"|X =="Bangladesh"|X=="Nepal") |>
  rename(	percentage_individuals_using_internet = "Value",
          country = "X",
          year = "Year") |>
  select(country, percentage_individuals_using_internet, year )
  internet_usage
  
```



##cleaning data
```{r}
education_graph <- southasian_countries_edu |>
  select(country , year, series, value) |>
  pivot_wider(names_from = series,
        values_from = value)
education_graph
```

##joining the main datasets without confounders

```{r}
joined_data <- southasian_countries |>
  left_join(education_graph , by = c("country", "year")) |>
  rename(primary_education_ratio = "Ratio of girls to boys in primary education",
         secondary_eductation_ratio = "Ratio of girls to boys in secondary education",
         uppersecondary_education_ratio = "Ratio of girls to boys in upper secondary education",
        women_seat_percent = "value") |>
  select(country, year, women_seat_percent, primary_education_ratio, secondary_eductation_ratio, uppersecondary_education_ratio )
joined_data
```

##running regression
```{r}
regression <- lm(women_seat_percent ~ primary_education_ratio , data = joined_data) 

regression2 <- lm(women_seat_percent ~ primary_education_ratio + secondary_eductation_ratio , data = joined_data) 

regression3 <- lm(women_seat_percent ~ primary_education_ratio + secondary_eductation_ratio + uppersecondary_education_ratio , data = joined_data) 

```

##tidying regression
```{r}
library(broom)

tidy(regression) |>
  knitr::kable()
```

##presenting regression
```{r}
library(modelsummary)

var_labels <- c(
  "primary_education_ratio" = "new name"
) ##u can change names here


modelsummary(list(regression, regression2, regression3),
              gof_map = c("nobs", "r.squared", "adj.r.squared"),
             coef_map = var_labels) ##this will run and pass the new name and make table more organized
```
##Result section 1




##joining confounder 1

```{r}
confounders <- joined_data |>
  left_join(GDP, by = c("country", "year"))
confounders
```
##joining confounder 2

```{r}
possible_confounders <- confounders |>
  left_join(internet_usage, by = c("country", "year"))
possible_confounders
```

##running regression with confounders
```{r}
regression4 <- lm(women_seat_percent ~ primary_education_ratio + secondary_eductation_ratio + uppersecondary_education_ratio + gdp_per_capita + percentage_individuals_using_internet, data = possible_confounders)
summary(regression4)
```

##presenting final regression
```{r}
modelsummary(regression4)
```
##Result section 2






##Conclusion

Confounders = reservation quota, terrorism, public expenditur on education, internet usage, GDP per capita, general liberal mindset has imcreased

question: should i remove the NA data?, SE kam hoga toh better haina?
